---
title: "simpleRegress"
author: "Eduardo Noriega"
date: "Fall 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
BEARS <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
BEARS$SEX<- as.factor(BEARS$SEX)
BEARS$MONTH<- as.factor(BEARS$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=BEARS)
plot(LENGTH~NECK,data=BEARS)
abline(simp)
summary.lm(simp)
```

## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).

```{r}
simp2 <- lm(LENGTH~HEADWTH,data=BEARS)
plot(LENGTH~HEADWTH,data=BEARS)
abline(simp2)
summary.lm(simp2)
  
```

The adjusted R-squared is 0.53, which is smaller than using the 'neck'.

```{r}
simp3 <- lm(LENGTH~WEIGHT,data=BEARS)
plot(LENGTH~WEIGHT,data=BEARS)
abline(simp3)
summary.lm(simp3)
  
```
 
The adjusted R-squared is 0.74, which is about the same as the 'neck'.

```{r}
simp4 <- lm(LENGTH~CHEST,data=BEARS)
plot(LENGTH~CHEST,data=BEARS)
abline(simp4)
summary.lm(simp4)
  
```

The adjusted R-squared is 0.78, which is bigger than the 'neck'.

```{r}
simp5 <- lm(LENGTH~HEADLEN,data=BEARS)
plot(LENGTH~HEADLEN,data=BEARS)
abline(simp5)
summary.lm(simp5)
  
```


The adjusted R-squared is 0.84 for HEADLEN, which is bigger than the 'neck' and 'chest'.  


```{r}
simp6 <- lm(LENGTH~as.numeric(SEX),data=BEARS)
plot(LENGTH~as.numeric(SEX),data=BEARS)
abline(simp6)
summary.lm(simp6)
  
```
This suggested R-squared is too small to be considered for 'sex' of the bear.  

```{r}
simp7 <- lm(LENGTH~as.numeric(MONTH),data=BEARS)
plot(LENGTH~as.numeric(MONTH),data=BEARS)
abline(simp7)
summary.lm(simp7)
  
```
This suggested R-squared is too small to be considered for 'month'.


















